---
title: "IS 605 - Assignment 1"
author: "Dan Fanelli"
date: "February 5, 2016"
output: html_document
---

## Problem Set 1:

Setup u and v:

```{r}
u <- c(0.5, 0.5)
v <- c(3,-4)
```

(1) Calculate the dot product u.v where u = [0.5; 0.5] and v = [3; −4]

```{r}
u.dot.v <- u %*% v
u.dot.v
```

(2) What are the lengths of u and v? Please note that the mathematical notion of the
length of a vector is not the same as a computer science definition.

```{r}
u.length <- length(u)
u.length

v.length <- length(v)
v.length
```

(3) What is the linear combination: 3u − 2v?

```{r}
linear.combination <- (3 * u) - (2 * v)
linear.combination
```

(4) What is the angle between u and v

(help from: http://stackoverflow.com/questions/1897704/angle-between-two-vectors-in-r)

```{r}
angle.between.u.and.v <- acos( sum(u*v) / ( sqrt(sum(u * u)) * sqrt(sum(v * v)) ) )
angle.between.u.and.v
```

## Problem Set 2:

A bit of help from: http://stackoverflow.com/questions/16044377/how-to-do-gaussian-elimination-in-r-do-not-use-solve

```{r}

solveWithElimination <- function(matrixRow1, matrixRow2, matrixRow3, matrixSolutionVector){
  A <- matrix(c(matrixRow1,matrixRow2,matrixRow3),byrow=T,nrow=3,ncol=3)
  b <- matrix(matrixSolutionVector,nrow=3,ncol=1)
  p <- nrow(A)
  (U.pls <- cbind(A,b))
  
  U.pls[1,] <- U.pls[1,]/U.pls[1,1]
  
  i <- 2
  while (i < p+1) {
   j <- i
   while (j < p+1) {
    U.pls[j, ] <- U.pls[j, ] - U.pls[i-1, ] * U.pls[j, i-1]
    j <- j+1
   }
   while (U.pls[i,i] == 0) {
    U.pls <- rbind(U.pls[-i,],U.pls[i,])
   }
   U.pls[i,] <- U.pls[i,]/U.pls[i,i]
   i <- i+1
  }
  for (i in p:2){
   for (j in i:2-1) {
    U.pls[j, ] <- U.pls[j, ] - U.pls[i, ] * U.pls[j, i]
   }
  }
  U.pls
}

```

Now invoke the function with values from stack overflow:

```{r}
matrixRow1 <- c(2,-5,4)
matrixRow2 <- c(1,-2.5,1)
matrixRow3 <- c(1,-4,6)
matrixSolutionVector <- c(-3,5,10)
  
solveWithElimination(matrixRow1, matrixRow2, matrixRow3, matrixSolutionVector)
```

Looks good to me, so now with homework values:

```{r}
matrixRow1 <- c(1,1,3)
matrixRow2 <- c(2,-1,5)
matrixRow3 <- c(-1,-2,4)
matrixSolutionVector <- c(1,2,6)
  
solveWithElimination(matrixRow1, matrixRow2, matrixRow3, matrixSolutionVector)
```
