---
title: "IS 605 - Assignment 4"
author: "Dan Fanelli"
date: "February 24, 2016"
output: pdf_document
---

## Problem Set 1: 

__Given a 3 × 3 matrix A, write code in R to compute X = AAT and Y = ATA.__
```{r}
A <- matrix(c(-1,1,0,1,3,-2,5,0,1), ncol=3)
X <- A %*% t(A)
Y <- t(A) %*% A
```

__EIGEN: "Then, compute the eigenvalues and eigenvectors of X and Y using the built-in commans in R."__
```{r}
eigen(X)$vectors
eigen(Y)$vectors
eigen(X)$values
eigen(Y)$values
```

__SVD: "Then, compute the left-singular, singular values, and right-singular vectors of A using the svd command"__
```{r}
svd_obj <- svd(A)
svd_obj$u
svd_obj$d
svd_obj$v
```

__"Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y"__
```{r}
# svd u and eigen x are related:
svd_obj$u
eigen(X)$vectors

# col 1: negatives of each other (adding yields zero)
round(svd_obj$u[,1] + eigen(X)$vectors[,1], digits = 4)
# col 2: negatives of each other (adding yields zero)
round(svd_obj$u[,2] + eigen(X)$vectors[,2], digits = 4)
# col 3: equal to each other (subtractinging yields zero)
round(svd_obj$u[,3] - eigen(X)$vectors[,3], digits = 4)

# svd v and eigen y are related:
svd_obj$v
eigen(Y)$vectors

# col 1: negatives of each other (adding yields zero)
round(svd_obj$v[,1] + eigen(Y)$vectors[,1], digits = 4)
# col 2: equal to each other (subtractinging yields zero)
round(svd_obj$v[,2] - eigen(Y)$vectors[,2], digits = 4)
# col 3: negatives of each other (adding yields zero)
round(svd_obj$v[,3] + eigen(Y)$vectors[,3], digits = 4)

```

## Problem Set 2:

__Write function to computer matrix inverse using co-factors:__

```{r warning=FALSE}
library(pracma)

get_co_factor <- function(subM, r, c){
  sign <- (-1)^(r+c)
  det <- det(subM)
  return (sign * det)
}

get_co_factor_matrix <- function(M){
  cofactor_M <- M;
  for(c in 1:col(M)){
    for(r in 1:row(M)){
      cofactor_M[r,c] <- get_co_factor(M,r,c)
    }
  }
  return (cofactor_M)
}

myinverse <- function(A){
  cofactor_M <- get_co_factor_matrix(A)
  transpose_M <- t(cofactor_M)
  return (transpose_M / det(A))
}

in1 <- matrix(c(4,3,3,2), nrow=2)
out1 <- myinverse(in1)
rref(in1 %*% out1)

in2 <- matrix(c(-24,20,-5,18,-15,4,5,-4,1), nrow=3)
out2 <- myinverse(in2)
rref(in2 %*% out2)

in3 <- matrix(c(2,14,7,46,5,24,3,19,7,63,5,4,31,2,35,26,17,8,19,16,75), nrow=5)
out3 <- myinverse(in3)
rref(in3 %*% out3)
```

