---
title: "IS 605 - Assignment 5"
author: "Dan Fanelli"
date: "February 29, 2016"
output: pdf_document
---

## Problem Set 1:

(1) Consider the unsolvable system Ax = b

### Write R markdown script to compute t(A)*A and t(A)*b

* First, a function for doing least squares:

```{r warning=FALSE}
# the least squares function:
run_least_squares <- function(A, b, show_data) {
  cat("---- run_least_squares start: \n")
  at_a <- t(A) %*% A
  at_b <- t(A) %*% b
  ### Solve for ^x in R using the above two computed matrices.
  x_star <- solve(at_a, at_b)
  ### What is the squared error of this solution?
  a_times_x_star <- A %*% x_star
  a_times_x_star_minus_b <- a_times_x_star - b
  length <- sqrt(sum(a_times_x_star_minus_b^2))
  the_error <- b - A %*% x_star

  if(show_data){
    max_2_show <- 10
    cat("A: ", head(A, n = max_2_show), "\n")
    cat("b: ", head(b, n = max_2_show), "\n")
    cat("at: ", head(t(A), n = max_2_show), "\n")
    cat("at_a: ", at_a[1:max_2_show], "\n")
    cat("at_b: ", at_b, "\n")
    cat("x_star: ", x_star, "\n")
    cat("(if x_star is 0s then the the following is really not needed...) \n")
    cat("a_times_x_star: ", a_times_x_star, "\n")
    cat("a_times_x_star_minus_b: ", a_times_x_star_minus_b, "\n")
    cat("length = ", length, "\n")
    cat("the_error = ", the_error, "\n")
    cat("---- run_least_squares finished. ---- \n")
  }else{
    cat("x_star: ", x_star, "\n")
  }
  
  return (x_star)
}

eval_root_mean_squared_error <- function(observed, predicted){
  root_mean_squared_error <- sqrt(mean((observed-predicted)^2))
  cat("root_mean_squared_error: ", root_mean_squared_error, "\n")
  observed_mean <- mean(observed)
  cat("observed_mean: ", observed_mean, "\n")
  rmse_div_by_observed_mean <- root_mean_squared_error / observed_mean
  cat("rmse_div_by_observed_mean: ", rmse_div_by_observed_mean)
}
```

### Run the script for the 2 equations:

```{r warning=FALSE}
A <- matrix(c(1,1,1,1,0,1,3,4), ncol=2)

# b, the first one that requires least squares (has no solution)
soln_b <- matrix(c(0,8,8,20), ncol=1)
soln_b_x_star <- run_least_squares(A,soln_b,TRUE)
soln_b_x_star

# p, the second one that does not require least squares (has a solution)
soln_p <- matrix(c(1,5,13,17), ncol=1)
soln_p_x_star <- run_least_squares(A,soln_p,TRUE)
soln_p_x_star

```

### Show that the error e = b - p = [-1; 3;-5; 3]

```{r warning=FALSE}
e <- soln_b - soln_p
e
eval_root_mean_squared_error(soln_b, soln_p)
```

### Show that the error e is orthogonal to p and to each of the columns of A.

```{r warning=FALSE}
#orthogonal means dot product = 0
sum(e*soln_p)
sum(e*A[,1])
sum(e*A[,2])
```

## Problem Set 2:

1. mpg:           continuous
2. cylinders:     multi-valued discrete
3. displacement:  continuous
4. horsepower:    continuous
5. weight:        continuous
6. acceleration:  continuous
7. model year:    multi-valued discrete
8. origin:        multi-valued discrete
9. car name:      string (unique for each instance)

### "Write an R markdown script that takes in the auto-mpg data, extracts an A matrix from the first 4 columns and b vector from the fifth (mpg) column."

* model mpg as a function of displacement, horsepower, weight, and acceleration
* read the data and show a quick summary/histogram of the target variable b
* strip all but the 4 important columns
* strip all the bad rows

```{r warning=FALSE}
data <- read.table("auto-mpg.data", sep="")
names(data) <- c("mpg", "cylinders", "displacement", "horsepower", "weight", "acceleration", "model_year", "origin", "car_name")
b <- as.matrix(data[,c("mpg")])
summary(b)
hist(b, xlim=c(-5,55), main="Distribution of b (mpg) from input data")
data <- data[,c("displacement","horsepower","weight","acceleration")]
#head(data)
data <- subset(data, displacement != '?' && horsepower != '?' && weight != '?' && acceleration != '?')
# convert to a matrix and show head
A <- data.matrix(data[1:4])
head(A)

```

### "Using the least squares approach, your code should compute the best fitting solution. That is, find the best fitting equation that expresses mpg in terms of the other 4 variables"

* So, using the function from above, without outputting all the data, let's find the equation __co-efficeints__ and take a peek at the __variable data__:

```{r warning=FALSE}
b_x_star_ps2 <- run_least_squares(A,b,TRUE)
b_x_star_ps2

displacement_coefficient <- b_x_star_ps2[1]
horsepower_coefficient <- b_x_star_ps2[2]
weight_coefficient <- b_x_star_ps2[3]
acceleration_coefficient <- b_x_star_ps2[4]

displacements <- A[,c("displacement")]
horsepowers <- A[,c("horsepower")]
weights <- A[,c("weight")]
accelerations <- A[,c("acceleration")]

head(displacements)
head(horsepowers)
head(weights)
head(accelerations)

```

* So the best fit least-squares __equation__ is: 

* __mpg = (-0.011258 * displacement) + (0.065595 * horsepower) + (-0.0010995 * weight) + (1.613 * acceleration)__

* Let's calculate a collection of mpgs using the above __equation__:

```{r warning=FALSE}

mpgs_from_equation <- (displacement_coefficient*displacements) + (horsepower_coefficient*horsepowers) + (weight_coefficient*weights) + (acceleration_coefficient*accelerations)

summary(mpgs_from_equation)
hist(mpgs_from_equation, xlim=c(-5,55), main="Distribution of calculated equation values (mpg) from input data")

```

### "Finally, calculate the fitting error between the predicted mpg of your model and the actual mpg."

* So the original was slightly skewed right, this is skewed left...and this seems to be represented in the summary and histograms of the b * x_star, where the error values are weighted towards the negative side...

```{r warning=FALSE}
eval_root_mean_squared_error(b, mpgs_from_equation)
```

