---
title: "DA 605 - Assignment 12"
author: "Dan Fanelli"
output: pdf_document
---

## BIAS VARIANCE TRADEOFF IN R

* Use the stats and boot libraries
* perform a cross-validation experiment to observe the bias variance tradeoff. 
* You'll use the auto data set from previous assignments. 
* This dataset has 392 observations across 5 variables. 
* We want to:
    + fit a polynomial model of various degrees using the glm function in R
    + measure the cross validation error using cv.glm function.

```{r message=FALSE, warning=FALSE}
library(boot)
library(stats)
library(knitr)

mpg_df <- read.table("auto-mpg.data", sep="")
# displacement, horsepower, weight, acceleration, mpg
names(mpg_df) <- c("disp", "hp", "wt", "acc", "mpg")
kable(head(mpg_df, n=10), align = "l")

```

* Fit various polynomial models to compute mpg as a function of the other four variables using glm function:
    + acceleration
    + weight
    + horsepower
    + displacement
* For example, the following will fit a 2nd degree polynomial function between mpg and the remaining 4 variables and perform 5 iterations of cross-validations.
    + glm.fit=glm(mpg~poly(disp+hp+wt+acc,2), data=auto) 
    + cv.err5[2] = cv.glm(auto,glm.fit,K=5)$delta[1] 
* This result will be stored in a cv.err5 array. 
* cv.glm returns the estimated cross validation error and its adjusted value in a variable called delta. * Please see the help on cv.glm to see more information.

* Once you have fit the various polynomials from degree 1 to 8, you can plot the crossvalidation error function as degree=1:8, plot(degree,cv.err5,type='b')

* For you assignment, please create an R-markdown document where you load the auto data set, perform the polynomial fit and then plot the resulting 5 fold cross validation curve. Your output should show the characteristic U-shape illustrating the tradeoff between bias and variance.

```{r message=FALSE, warning=FALSE}
degree_errors = c()
degree_values = c(1:8)
for(degree in degree_values)
{
  glm.fit=glm(mpg~poly(disp+hp+wt+acc,2), data=mpg_df)
  deg_err <- cv.glm(mpg_df, glm.fit, K=5)$delta[1]
  degree_errors[degree] <- deg_err
}

plot(degree_values, degree_errors, type='b')

```
