---
title: "DA 605 - Assignment 11"
author: "Dan Fanelli"
output: pdf_document
---

## LINEAR REGRESSION IN R

* Your submission should include the final linear fits, and their corresponding significance levels. 
* In addition, you should clearly state what you concluded from looking at the fit and their significance levels.

```{r message=FALSE, warning=FALSE}
library(knitr)
library(ggplot2)
library(rpart)
```

*****
### Age and Max HR:
*****

__We hear that MaxHR = 220 - Age__

```{r message=FALSE, warning=FALSE}
age <- c(18,23,25,35,65,54,34,56,72,19,23,42,18,39,37)
maxHR <- c(202,186,187,180,156,169,174,172,153,199,193,174,198,183,178)
hr_age_df <- data.frame(age, maxHR)
summary(hr_age_df)
kable(hr_age_df, align = "l")
```

Using R's lm function:

1. Perform regression analysis

```{r message=FALSE, warning=FALSE}
hr_age_model <- lm(maxHR ~ age)
summary(hr_age_model)
```

2. Measure the signicance of the independent variables

```{r message=FALSE, warning=FALSE}
cor(age, maxHR)
```

___The independent variable 'age' has a *** next to it, so its significance level is basically 0.___

3. What is the resulting equation?

```{r message=FALSE, warning=FALSE}
HR = 210.04846 + (-0.79773)*age
```

4. Is the effect of Age on Max HR significant?

___Its basically 0, so not significant.___

5. What is the signicance level? 

___The significant level is *** which is basically zero.___

6. Please also plot the fitted relationship between Max HR and Age.

```{r message=FALSE, warning=FALSE}
ggplot(data = hr_age_df, aes(x = age, y = maxHR)) + geom_smooth(method = "lm", se=FALSE, color="blue", formula = y ~ x) + geom_point() + geom_abline(intercept = 220, slope = -1)
```

*****
### Auto Data:
*****

Using the Auto data set from Assignment 5 (also attached here) perform a Linear Regression analysis using mpg as the dependent variable and the other 4 (displacement, horsepower, weight, acceleration) as independent variables. 

```{r message=FALSE, warning=FALSE}
mpg_df <- read.table("auto-mpg.data", sep="")
names(mpg_df) <- c("mpg", "cylinders", "displacement", "horsepower", "weight", "acceleration", "model_year", "origin", "car_name")

mpg_df <- mpg_df[,c("mpg","displacement","horsepower","weight","acceleration")]
mpg_df <- subset(mpg_df, displacement != '?' && horsepower != '?' && weight != '?' && acceleration != '?')

# it was showing horsepower as categorical, so giving strange results
mpg_df <- transform(mpg_df, horsepower = as.numeric(horsepower))

kable(head(mpg_df, n=10), align = "l")

```

1. What is the final linear regression fit equation? 

```{r message=FALSE, warning=FALSE}
mpg_model <- lm(mpg ~ mpg_df$displacement + mpg_df$horsepower + mpg_df$weight + mpg_df$acceleration, data = mpg_df)

summary(mpg_model)

```

___So the final linear regression fit equation is:___

mpg = 40 - (displacement * -0.0106291) + (horsepower * 0.0047774) + (weight * -0.0061405) + (acceleration * 0.1722165)

2. Which of the 4 independent variables have a significant impact on mpg? 

Based on the significance codes, only weight seems to have zero impact on mpg.

3. What are their corresponding significance levels? 

* displacement: 0.1041
* horsepower: 0.5633
* weight: (zero...)
* acceleration: 0.0785

4. What are the standard errors on each of the coeficients? 

* displacement: 1.9966258
* horsepower: 0.0065254
* weight: 0.0007449
* acceleration: 0.0976340

Please perform this experiment in two ways. 

1. First take any random 40 data points from the entire auto data sample and perform the linear regression fit and measure the 95% confidence intervals. 

```{r message=FALSE, warning=FALSE}
mpg_df_40_sample <- mpg_df[sample(nrow(mpg_df), 40), ]
mpg_df_40_sample_fit <- lm(mpg ~ displacement + horsepower + weight + acceleration, data=mpg_df_40_sample)
summary(mpg_df_40_sample_fit)
confint(mpg_df_40_sample_fit, level=0.95)
plot(mpg_df_40_sample_fit)
cor(mpg_df_40_sample$displacement, mpg_df_40_sample$mpg)
cor(mpg_df_40_sample$horsepower, mpg_df_40_sample$mpg)
cor(mpg_df_40_sample$weight, mpg_df_40_sample$mpg)
cor(mpg_df_40_sample$acceleration, mpg_df_40_sample$mpg)
```

2. Then, take the entire data set (all 392 points) and perform linear regression and measure the 95% confidence intervals. 

```{r message=FALSE, warning=FALSE}
mpg_df_all <- mpg_df
mpg_df_all_fit <- lm(mpg ~ displacement + horsepower + weight + acceleration, data=mpg_df_all)
summary(mpg_df_all_fit)
confint(mpg_df_all_fit, level=0.95)
plot(mpg_df_all_fit)
cor(mpg_df_all$displacement, mpg_df_all$mpg)
cor(mpg_df_all$horsepower, mpg_df_all$mpg)
cor(mpg_df_all$weight, mpg_df_all$mpg)
cor(mpg_df_all$acceleration, mpg_df_all$mpg)
```

3. Please report the resulting fit equation, their significance values and confidence intervals for each of the two runs.

(done, see above...)

# Some Help Analysis / Debugging:

from: http://www.statmethods.net/stats/regression.html

```{r message=FALSE, warning=FALSE}
# Other useful functions 
coefficients(mpg_df_all_fit) # model coefficients
confint(mpg_df_all_fit, level=0.95) # CIs for model parameters 
#fitted(mpg_df_all_fit) # predicted values
#residuals(mpg_df_all_fit) # residuals
anova(mpg_df_all_fit) # anova table 
vcov(mpg_df_all_fit) # covariance matrix for model parameters 
#influence(mpg_df_all_fit) # regression diagnostics

###########
ggplot(data = mpg_df, aes(y = mpg, x = displacement)) +     geom_smooth(method = "lm", color="green", formula = y ~ x) + geom_point()

ggplot(mpg_df, aes(y=mpg, x=horsepower)) + geom_smooth(method = "lm", color="blue", formula = y ~ x) + geom_point()

ggplot(mpg_df, aes(y=mpg, x=weight)) + geom_smooth(method = "lm", color="orange", formula = y ~ x) + geom_point()

ggplot(mpg_df, aes(y=mpg, x=acceleration)) + geom_smooth(method = "lm", color="red", formula = y ~ x) + geom_point()

```